---
title: "Distribuições de Probabilidade"
author: "Jaqueline V Moreira"
date: "2024-06-12"
output: powerpoint_presentation
---

```{r "Loading packages", include=FALSE}
library(tidyverse)
library(ggpubr) #organizar e exibir vários gráficos juntos
library(ggalt) #ggplots com linhas mais suaves
library(dplyr)
library (magrittr)# compound assignment pipe-operator - %<>%
library (patchwork) # apresentar mais de um gráfico junto

# library(scales) #converter escala decimal para percentual
# library(queueing) # modelo de enfileiramento

```

```{r "color palettes", include=FALSE}
color_binomial = c("#ab9b96", "purple")

color_poisson = c("#DCAB6B","darkorange","#F28F3B","#C0BF5F","#C8553D",
                  "#BB7E5D","#BA6E6E","#8C705F", "#DE5466")

color_expo = c( "#FEA82F", "#FF2E00", "#FF4500", "#FF8C00")



color_urban= c("#2E2E3D", "#BFBFC0", "#C0BF5F", "#55556F", "#0C6291")

Color_violet = c("#624763", "#F9B3D1", "#2F1847", "#C62E65", "#D63AF9", "#A63446", "#7E1946", "#FF0022")

color_candy = c("#26547C", "#EF476F", "#FFD166", "#06D6A0", "#FF0022")

```

# Binomial Distribution

According to Statistics Canada life tables, the probability a randomly selected 90 year-old Canadian male survives for at least another year is approximately 0.82.
If twenty 90-year-old Canadian males are randomly selected, 

a) what is the probability exactly 18 survives for at least another year?
b) what is the probability that last than 11 survives for another year?

Successes: The man survives for at least on year

Failure: The man dies within one year
Since the sampling of men are random, it’s reasonable to consider them independent

Let x represent the number of men that survive for at least one year
n = 20
p = 0.82



```{r "Binomial data frame", echo=FALSE, message=FALSE, warning=FALSE}

n = 20
p = 0.82

success <- 0:21
probability <- rep(NA,length(success))

for (i in success){
 
    probability[i] =  dbinom(i, size = n, prob = p)}

binomial_df = data.frame("Successes" = success, "Probability" = probability)
```

```{r "Binomial exercise 1", echo=FALSE, message=FALSE, warning=FALSE}
x = 18
n = 20
p = 0.82

asw = dbinom(x = x, size = n, prob = p)

paste0("P(x = 18) = ", round(asw,5))

  # filtrando dados para destaque no gráfico
  # é necessário criar uma variavel que indique quais linhas do data frame devem ser 
  # destacadas. Para isso, utiliza-se a função mutate para criar a variável e a função     # case_when para selecionar as linhas de interesse.

 binomial_df = binomial_df %>% mutate(x_igual_18 = case_when(Successes==17 ~ TRUE,TRUE ~ FALSE))

```

```{r "Binomial exercise 1 Plot", echo=FALSE, message=FALSE, warning=FALSE}

binomial_plot_1 = binomial_df %>% ggplot(mapping = (aes(x=Successes))) +
  
  # Construindo gráficos
  geom_col(mapping = (aes(y = probability, fill = x_igual_18)))+
  
  # definindo cores
  scale_fill_manual(values = color_binomial)+

  # definindo Titulos
  labs(title = "x~Bin(20,0.82)", 
       subtitle = "Binomial Distribution with n = 20 and p = 0.82",
       y = "P(x)",
       x= "Number of successes") + 

       
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
    
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

binomial_plot_1
```

```{r "Binomial exercise 2", echo=FALSE, message=FALSE, warning=FALSE}

  # filtrando dados para destaque no gráfico
  # é necessário criar uma variavel que indique quais linhas do data frame devem ser 
  # destacadas. Para isso, utiliza-se a função mutate para criar a variável e a função     # case_when para selecionar as linhas de interesse.

 binomial_df = binomial_df %>% mutate(x_menor_15 = case_when(Successes<15 ~ TRUE,TRUE ~ FALSE))


  # Gerando probabilidade acumulada de acordo com o destaque anterior

  sum_bin = 0
  
  for (i in binomial_df$Successes){
    if (binomial_df$x_menor_15[i+1] == TRUE){
      sum_bin = sum_bin+binomial_df$Probability[i+1] 
    }}

paste0("P(x<15) = ", round(sum_bin,4))

```

```{r "Binomial exercise 2 Plot", echo=FALSE, message=FALSE, warning=FALSE}

binomial_plot_2 = binomial_df %>% ggplot(mapping = (aes(x=Successes))) +
  
  # Construindo gráficos
  geom_col(mapping = (aes(y = probability, fill = x_menor_15)))+
  
  # definindo cores
  scale_fill_manual(values = color_binomial )+
  
  # definindo Titulos
  labs(title = "x~Bin(20,0.82)", 
       subtitle = "Binomial Distribution with n = 20 and p = 0.82",
       y = "P(x)",
       x= "Number of successes") + 

  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
    
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

binomial_plot_2
```


# Poisson Distribution

```{r "Poisson lambda calcs", message=FALSE, warning=FALSE, include=FALSE}

lambda = c(0.5,1,1.5,2,4,5)
x_axis = c(0:35)


lambda_0.5 =   rep(NA,length(x_axis))
lambda_1   =   rep(NA,length(x_axis))
lambda_1.5 =   rep(NA,length(x_axis))
lambda_2   =   rep(NA,length(x_axis))
lambda_3   =   rep(NA,length(x_axis))
lambda_5   =   rep(NA,length(x_axis))
lambda_10   =  rep(NA,length(x_axis))
lambda_15   =  rep(NA,length(x_axis))

for( i in x_axis ) {
  lambda_0.5[i+1] =  dpois(x=i, lambda = 0.5)
  lambda_1[i+1]   =  dpois(x=i, lambda = 1)
  lambda_1.5[i+1] =  dpois(x=i, lambda = 1.5)
  lambda_2[i+1]   =  dpois(x=i, lambda = 2)
  lambda_3[i+1]   =  dpois(x=i, lambda = 3)
  lambda_5[i+1]   =  dpois(x=i, lambda = 5)
  lambda_10[i+1]  =  dpois(x=i, lambda = 10)
  lambda_15[i+1]  =  dpois(x=i, lambda = 15)
  }


poisson_data = data.frame("x" = x_axis, 
                          "lam_0.5" =lambda_0.5, 
                          "lam_1" = lambda_1, 
                          "lam_1.5" = lambda_1.5, 
                          "lam_2" = lambda_2, 
                          "lam_3" = lambda_3, 
                          "lam_5" = lambda_5,
                          "lam_10" = lambda_10,
                          "lam_15" = lambda_15)

poisson_data

```

```{r "Poisson lambda plots", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

poisson_plots = poisson_data %>% ggplot(mapping = (aes(x=x))) +
  
  # Construindo gráficos
  geom_line(mapping = (aes(y = lam_1, color = "0.5")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_1, color = "0.5")))+
  
  geom_line(mapping = (aes(y = lam_1.5, color = "1.5")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_1.5, color = "1.5")))+
  
  geom_line(mapping = (aes(y = lam_2, color = "2.0")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_2, color = "2.0")))+
  
  geom_line(mapping = (aes(y = lam_3, color = "3.0")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_3, color = "3.0")))+
  
  geom_line(mapping = (aes(y = lam_5, color = "5.0")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_5, color = "5.0")))+
  
  geom_line(mapping = (aes(y = lam_10, color = "10.0")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_10, color = "10.0")))+
  
  geom_line(mapping = (aes(y = lam_15, color = "15.0")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_15, color = "15.0")))+
  
  # # definindo cores
  scale_color_manual(values=color_poisson)+
  
  # definindo Titulos
  labs(title = "Poisson Distribution", 
       subtitle = "with n = 35 and Lambda from 0.5 to 5",
       y = "P(x)",
       x= "Number of successes for time interval",
       color = "λ = ")+ 
 
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
    
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

poisson_plots



```


Exclusive Vines import Argentinian wine into australia. They've begun advertising on Facebook to direct traffic to their website where customers can order wine online. The number of click-through sales from the ad is Poisson distributed with a mean of 12 click-through sales per day.

μ = 12 click-through sales per day
E(x) = μ = λ

Find the probability of getting:

a) exaclty 8 click-through sales in the first day
b) at least 10 click-through sales in the first day


```{r "ex lambda 12 calc", echo=FALSE, message=FALSE, warning=FALSE}

x_axis = c(0:35)
prob_lambda = rep(NA,length(x_axis))

for( i in x_axis ) {
  prob_lambda[i+1] =  dpois(x=i, lambda = 12)}

lambda_12 = data.frame("x_axis" = x_axis, "prob_lambda" = prob_lambda)

```

```{r "ex lambda 12 filtros de destaque", echo=FALSE, message=FALSE, warning=FALSE}

  # filtrando dados para destaque
  # é necessário criar uma variavel que indique quais linhas do data frame devem ser 
  # destacadas. Para isso, utiliza-se a função mutate para criar a variável e a função     # case_when para selecionar as linhas de interesse.

  lambda_12_df <- lambda_12 %>% mutate(x_maior_10 = case_when(x_axis>=10 ~ TRUE, 
                                                                  TRUE ~ FALSE),
                                        x_igual_8 = case_when(x_axis==8 ~ TRUE, 
                                                                  TRUE ~ FALSE))

  # # o ggplot só realiza a sobreposição de gráficos quando todos possuem a mesma
  # # quantidade de valores x ou quando se tratar de um único ponto ou coluna.


  # Gerando probabilidade acumulada de acordo com o destaque anterior

  sum_poi_menor_10 = 0
  
  for (i in lambda_12_df$x_axis){
    if (lambda_12_df$x_maior_10[i+1] == FALSE){
      sum_poi_menor_10 = sum_poi_menor_10+lambda_12_df$prob_lambda[i+1] 
    }}
  
  sum_poi_maior_10 = 1-sum_poi_menor_10

paste0("P(x=8) = ", round(lambda_12_df$prob_lambda[9],4))
paste0("P(x>=10) = 1-P(x<10)")
paste0("P(x>=10) = ", round(sum_poi_maior_10,4))

```

```{r "ex lambda 12 plot", echo=FALSE, message=FALSE, warning=FALSE}
poisson_ex_plot <-lambda_12_df%>%ggplot(mapping = (aes(x = x_axis, y = prob_lambda))) +
  
  # Gerando gráfico
  geom_col(mapping = (aes()), fill = "#DCAB6B", width = 0.5)+

  # sobrepondo curva suavizada
  geom_xspline(mapping = (aes()),
                          color= "#6E0D25", 
                          spline_shape = -0.4)+
  
  # xpline suaviza gráficos de linha, o sline_shape indicado é de -0.4. 
  # quando sline_shape = 0 não há suavização
  
  geom_point(mapping = (aes()), 
                        color = "#6E0D25")+
 
  
  # definindo Titulos
  labs(title = "X~Poi(12)", 
       subtitle = "Poisson Distribution with λ = 12",
       y = "P(x)",
       x= "Number of successes for time interval") + 
 
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size=11),
    
        #legend
        legend.position= "none",
        legend.text = element_blank(),
        
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

poisson_ex_plot

```

a) exaclty 8 click-through sales in the first day

```{r "ex lambda 12 exacly 8 plot", echo=FALSE, message=FALSE, warning=FALSE}

poisson_ex_plot_2 <- lambda_12_df %>%  ggplot(mapping = (aes(x = x_axis))) +

  geom_col(mapping = (aes(x = x_axis,
                          y = prob_lambda,  
                          fill = x_igual_8)), #segmentar por variavel True ou False
                      width = 0.5)+
  
  # sobrepondo curva suavizada
  geom_xspline(mapping = (aes(y = prob_lambda)),
                          color= "#6E0D25", 
                          spline_shape = -0.4)+
  
  # xpline suaviza gráficos de linha, o sline_shape indicado é de -0.4. 
  # quando sline_shape = 0 não há suavização
  
  geom_point(mapping = (aes(y = prob_lambda)), 
                        color = "#6E0D25")+
  
  
  # definindo cores
  scale_fill_manual(values=color_poisson)+
  
  # definindo Titulos
  labs(title = "X~Poi(12)", 
       subtitle = "Poisson Distribution with λ = 12",
       y = "P(x)",
       x= "Number of successes for time interval",
       fill = "exacly 8 successes") + 
 
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size=11),
    
        #legend
        legend.position= "none",
        legend.text = element_blank(),
        
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

poisson_ex_plot_2


```

b) at least 10 click-through sales in the first day

```{r "ex lambda 12 at least 10 plot", echo=FALSE, message=FALSE, warning=FALSE}

poisson_ex_plot_3 <- lambda_12_df %>%  ggplot(mapping = (aes(x = x_axis))) +

  geom_col(mapping = (aes(x = x_axis,
                          y = prob_lambda,  
                          fill = x_maior_10)), #segmentar por variavel True ou False
                      width = 0.5)+
  
  # sobrepondo curva suavizada
  geom_xspline(mapping = (aes(y = prob_lambda)),
                          color= "#6E0D25", 
                          spline_shape = -0.4)+
  
  # xpline suaviza gráficos de linha, o sline_shape indicado é de -0.4. 
  # quando sline_shape = 0 não há suavização
  
  geom_point(mapping = (aes(y = prob_lambda)), 
                        color = "#6E0D25")+
  
  
  # definindo cores
  scale_fill_manual(values=color_poisson)+
  
  # definindo Titulos
  labs(title = "X~Poi(12)", 
       subtitle = "Poisson Distribution with λ = 12",
       y = "P(x)",
       x= "Number of successes for time interval",
       fill = "exacly 8 successes") + 
 
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size=11),
    
        #legend
        legend.position= "none",
        legend.text = element_blank(),
        
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

poisson_ex_plot_3

```

## Relationship between binomial and poisson distribution
```{r "binomial and poisson calc"}
p = 0.05 # 5%
n = c(900:3000)
x = 90

bin = rep(NA, length(n))
poi = rep(NA, length(n))


for (b in n) {
  bin[b-899] = dbinom(x=x,size = b, prob = p)
}

for (i in n) {
  poi[i-899] = dpois(x=x, lambda = i*p)
}

bin_poi = data.frame("x_axis" = n, "bin" = bin, "poi" = poi)

```

```{r "binomial and poisson plot"}

bin_poi_plots = bin_poi %>% ggplot(mapping = (aes(x=x_axis))) +
  
  # Construindo gráficos
  geom_line(mapping = (aes(y = bin, color = "Binomial")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = bin, color = "Binomial")))+
  
  geom_line(mapping = (aes(y = poi, color = "Poisson")), linewidth = 0.7, linetype = 2)+
  geom_point(mapping = (aes(y = poi, color = "Poisson")))+
  
  # # definindo cores
  scale_color_manual(values=color_poisson)+
  
  # definindo Titulos
  labs(title = "Binomial and Poisson Distribution", 
       subtitle = "n from 900 to 3000, p = 5% and Lambda = np",
       y = "P(x)",
       x= "Number of successes",
       color = "Distribuition")+ 
 
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
    
        #legend
        legend.position="top",
        legend.text = element_text(size=10),
        
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

bin_poi_plots

```

```{r "binomial and poisson difference"}
bin_x_90 = round(dbinom(x=x,size = 2500, prob = p),6)
poi_x_90 = round(dpois(x=x,lambda = 2500*p),6)


paste0("X~Bin(2500,0.05)")
paste0("P(X = 90) = ",bin_x_90 )
paste0(" and ")
paste0("X~Poi(2500*0.05)")
paste0("P(X = 90) = ",poi_x_90 )
paste0("Uma diferença de apenas ",round(poi_x_90 - bin_x_90,6)) 

```


### binomial and poisson example

Albinism is a rare genetic disorder that affects one in 20.000 europeans.
People with albinism produce little or none of the pigment melanin.
In a random sample of 1.000 europeans, what is the probability that exactly 2 have albnism?

In a random sampling the events can be considered independent. In this case each trial can be considered a bernoulli trial with

p = 1/20000

In a Binomial distribution:
n = 1000
x = 2

in a Poisson distribution:
Lambda = n*p
lambda = 1000*1/20000

```{r "binomial and poisson example"}
x = 2
p = 1/20000
n = 1000
lambda = n*p

bin_x_2 = round(dbinom(x=x,size = n, prob = p),6)
poi_x_2 = round(dpois(x=x,lambda = lambda),6)

paste0("In a Binomial distribution P(X = 2) = ", bin_x_2)
paste0("In a Poisson distribution P(X = 2) = ", poi_x_2)
paste0("a difference of ", round(poi_x_2 -bin_x_2,6))

```




# Exponential Distribution

```{r "Expo lambda calcs", echo=FALSE, message=FALSE, warning=FALSE}
lambda = c(1.5,3,6,12,18,24,30)
rate_expo = c(0.025,0.5,0.1,0.2,0.3,0.4,0.5)
x_axis = 0:35

lambda_1_5 =   rep(NA,length(x_axis))
lambda_3   =   rep(NA,length(x_axis))
lambda_6   =   rep(NA,length(x_axis))
lambda_12  =   rep(NA,length(x_axis))
lambda_18  =   rep(NA,length(x_axis))
lambda_24  =   rep(NA,length(x_axis))
lambda_30  =   rep(NA,length(x_axis))

for (i in x_axis) {
  lambda_1_5[i+1] =  dexp(x=i, rate = 0.025) # lambda de 1,5 por hora = 1.5/60min  = 0.025 por min
  lambda_3[i+1]   =  dexp(x=i, rate = 0.05)  # lambda de 3 por hora = 3/60min  = 0.05 por min
  lambda_6[i+1]   =  dexp(x=i, rate = 0.1)   # lambda de 6 por hora = 6/60min  = 0.10 por min
  lambda_12[i+1]  =  dexp(x=i, rate = 0.2)   # lambda de 12 por hora = 12/60min  = 0.2 por min
  lambda_18[i+1]  =  dexp(x=i, rate = 0.3)   # lambda de 18 por hora = 18/60min  = 0.3 por min
  lambda_24[i+1]  =  dexp(x=i, rate = 0.4)   # lambda de 24 por hora = 24/60min  = 0.4 por min
  lambda_30[i+1]  =  dexp(x=i, rate = 0.5)   # lambda de 30 por hora = 30/60min  = 0.5 por min
}

# o rate da distribuição exponencial deve ser descrito em termos da menor unidade de tempo
# por exemplo, a chegada de clientes condiz com uma distribuição poisso com lambda = 3 por hora
# dessa forma, a distribuição exponencial desses clientes será 3/60min ou 0.05


expo_data = data.frame("x" = x_axis, 
                        "lambda_1_5" =  lambda_1_5, 
                        "lambda_3"   =  lambda_3, 
                        "lambda_6"   =  lambda_6, 
                        "lambda_12"  =  lambda_12, 
                        "lambda-18"  =  lambda_18,
                        "lambda_24"  =  lambda_24, 
                        "lambda_30"  =  lambda_30)
                    

expo_data

```

```{r "Expo lambda plots", message=FALSE, warning=FALSE, paged.print=FALSE}

expo_plots = expo_data %>% ggplot(mapping = (aes(x=x_axis))) +
  
  # Construindo gráficos
  geom_line(mapping = (aes(y = lambda_1_5, color = "1.5")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_1_5, color = "1.5")))+
  
  geom_line(mapping = (aes(y = lambda_3, color = "3")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_3, color = "3")))+
  
  geom_line(mapping = (aes(y = lambda_6, color = "6")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_6, color = "6")))+

  geom_line(mapping = (aes(y = lambda_12, color = "12")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_12, color = "12")))+
  
  geom_line(mapping = (aes(y = lambda_18, color = "18")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_18, color = "18")))+
  
  geom_line(mapping = (aes(y = lambda_24, color = "24")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_24, color = "24")))+
  
  geom_line(mapping = (aes(y = lambda_30, color = "30")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_30, color = "30")))+
  
  # # definindo cores
  scale_color_manual(values=Color_violet)+
  
  # definindo Titulos
  labs(title = "Exponencial Distribution", 
       subtitle = "with n = 35 and Lambda from 1.5 to 30",
       y = "P(x)",
       x= "Time for next arrival",
       color = "λ = ")+ 
 
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
    
        #legend
        legend.position="top",
        legend.text = element_text(size=10),
        
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

expo_plots



```


Unique visitors arrive at a certain site by a Poisson distribution at an average rate of 3 per hour.

the average rate of an poisson distribution is lambda so in this case lambda is iqual to 3/hour

What is the probability that the next visitor arrives:

a) Within 10 mins
b) after 30 mins passes
c) between 40 and 60 mins

The questions references the time interval for the next arrival NOT THE AMOUNT OF VISITORS. In this case, the exponential distribution will be used.  

lambda of 3 means that 3 visitors arrive by hour or one visitor every 20 min on average.
To make the graph more intuitive, the scale will be build in terms of minutes.

3 per hour = 3/60 = 0.05 clientes per minute (rate in the graph)

```{r "Expo lambda = 3", echo=FALSE, message=FALSE, warning=FALSE}

lambda = 3
rate = 0.05
x_axis = 0:180 # minutos

expo_3 = rep(NA, length(x_axis))

for (i in x_axis){
  expo_3[i+1] = dexp(x=i, rate = rate) 
}

expo_3_df = data.frame("x_axis" = x_axis, "lambda_3" = expo_3)
expo_3_df

```

```{r "Expo lambda = 3 plot", warning=FALSE, paged.print=FALSE}

expo_3_plot = expo_3_df %>% ggplot(mapping = (aes(x=x_axis))) +
  
  # Construindo gráficos
  geom_line(mapping = (aes(y = lambda_3, color = "3")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_3, color = "3")))+
  # # definindo cores
  scale_color_manual(values=Color_violet)+
  
  # definindo Titulos
  labs(title = "Exponencial Distribution", 
       subtitle = "λ = 3 exponencial rate = 0.05",
       y = "P(x)",
       x= "Time for next arrival",
       color = "λ = ")+ 
 
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
    
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

expo_3_plot



```

### a) within 10 min

```{r "Expo lambda = 3 within 10 min", echo=FALSE, message=FALSE, warning=FALSE}

# filtrando dados para destaque no gráfico
# é necessário criar uma variavel que indique quais linhas do data frame devem ser 
# destacadas. Para isso, utiliza-se a função mutate para criar a variável e a função     
# case_when para selecionar as linhas de interesse.

 expo_3_df = expo_3_df %>% mutate(x_menor_15 = case_when(x_axis<=15 ~ TRUE,TRUE ~ FALSE))


# Gerando probabilidade acumulada de acordo com o destaque anterior

 sum_ex_1 = 0
  
  for (i in expo_3_df$x_axis){
    if (expo_3_df$x_menor_15[i+1] == TRUE){
      sum_ex_1 = sum_ex_1+expo_3_df$lambda_3[i+1] 
    }}
 
expo_menor_15 = pexp(q = 15, rate = 0.05)

paste0("P(x<15) = ", round(expo_menor_15,4))

```

```{r "Expo lambda = 3 within 10 min plot", warning=FALSE, paged.print=FALSE}

expo_3_plot = expo_3_df %>% ggplot(mapping = (aes(x=x_axis))) +
  
  # Construindo gráficos
  geom_line(mapping = (aes(y = lambda_3, color = "3")), linewidth = 0.9)+
  #geom_point(mapping = (aes(y = lambda_3, color = "3")))+
  geom_area(mapping = (aes(y = lambda_3, fill = x_menor_15)))+
  annotate("text", x=15, y = 0.04, label = "57,76%")+
  
  # # definindo cores
  scale_color_manual(values=Color_violet)+
  scale_fill_manual(values=Color_violet)+
  
  # definindo Titulos
  labs(title = "Exponencial Distribution", 
       subtitle = "λ = 3",
       y = "P(x)",
       x= "Time for next arrival",
       color = "λ = ")+ 
 
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
      
    
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

expo_3_plot



```

### b) after 30 min

```{r "Expo lambda = 3 after 30 min", echo=FALSE, message=FALSE, warning=FALSE}

# filtrando dados para destaque no gráfico
# é necessário criar uma variavel que indique quais linhas do data frame devem ser 
# destacadas. Para isso, utiliza-se a função mutate para criar a variável e a função     
# case_when para selecionar as linhas de interesse.

 expo_3_df = expo_3_df %>% mutate(x_maior_30 = case_when(x_axis>=30 ~ TRUE,TRUE ~ FALSE))


# Gerando probabilidade acumulada de acordo com o destaque anterior

expo_maior_30 = 1 - (pexp(q = 30, rate = 0.05))

paste0("P(x>30) = ", round(expo_maior_30,4))

```

```{r "Expo lambda = 3 after 30 min plot", warning=FALSE, paged.print=FALSE}

expo_3_plot = expo_3_df %>% ggplot(mapping = (aes(x=x_axis))) +
  
  # Construindo gráficos
  geom_line(mapping = (aes(y = lambda_3, color = "3")), linewidth = 0.9)+
  #geom_point(mapping = (aes(y = lambda_3, color = "3")))+
  geom_area(mapping = (aes(y = lambda_3, fill = x_maior_30)))+
  annotate("text", x=70, y = 0.007, label = round(expo_maior_30,4))+
  
  # # definindo cores
  scale_color_manual(values=Color_violet)+
  scale_fill_manual(values=Color_violet)+
  
  # definindo Titulos
  labs(title = "Exponencial Distribution", 
       subtitle = "λ = 3",
       y = "P(x)",
       x= "Time for next arrival",
       color = "λ = ")+ 
 
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
      
    
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

expo_3_plot



```


### b) between 40 and 60 min

```{r "Expo lambda = 3 between 40 and 60 min", echo=FALSE, message=FALSE, warning=FALSE}

# filtrando dados para destaque no gráfico
# é necessário criar uma variavel que indique quais linhas do data frame devem ser 
# destacadas. Para isso, utiliza-se a função mutate para criar a variável e a função     
# case_when para selecionar as linhas de interesse.

 expo_3_df = expo_3_df %>% mutate(x_menor_60 = case_when(x_axis<=60 ~ TRUE,TRUE ~ FALSE))
 expo_3_df = expo_3_df %>% mutate(x_menor_40 = case_when(x_axis<=40 ~ TRUE,TRUE ~ FALSE))


# Gerando probabilidade acumulada de acordo com o destaque anterior


expo_between_40_60 = (pexp(q = 60, rate = 0.05)) - (pexp(q = 40, rate = 0.05))

paste0("P(x>40 e x<60) = ", round(expo_between_40_60,4))

```

```{r "Expo lambda = 3 between 40 and 60 min plot", warning=FALSE, paged.print=FALSE}

expo_3_60_plot = expo_3_df %>% ggplot(mapping = (aes(x=x_axis))) +
  
  # Construindo gráficos
  #geom_line(mapping = (aes(y = lambda_3, color = "3")), linewidth = 0.9)+
  #geom_point(mapping = (aes(y = lambda_3, color = "3")))+
  geom_area(mapping = (aes(y = lambda_3, fill = x_menor_60)))+
  annotate("text", x=60, y = 0.025, label = round(pexp(60,rate = 0.05),4))+
  
  # # definindo cores
  scale_color_manual(values=color_urban)+
  scale_fill_manual(values=color_urban)+
  
  # definindo Titulos
  labs(title = "Exponencial Distribution", 
       subtitle = "λ = 3",
       y = "P(x)",
       x= "Time for next arrival",
       color = "λ = ")+ 
 
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
      
    
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        
        #axis
        axis.text.x = element_blank(),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

######################################

expo_3_40_plot = expo_3_df %>% ggplot(mapping = (aes(x=x_axis))) +
  
  # Construindo gráficos
  #geom_line(mapping = (aes(y = lambda_3, color = "3")), linewidth = 0.9)+
  #geom_point(mapping = (aes(y = lambda_3, color = "3")))+
  geom_area(mapping = (aes(y = lambda_3, fill = x_menor_40)))+
  annotate("text", x=40, y = 0.025, label = round(pexp(40,rate = 0.05),4))+
  
  # # definindo cores
  scale_color_manual(values=Color_violet)+
  scale_fill_manual(values=Color_violet)+
  
  # definindo Titulos
  labs(title = "Exponencial Distribution", 
       subtitle = "λ = 3",
       y = "P(x)",
       x= "Time for next arrival",
       color = "λ = ")+ 
 
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_blank(),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_blank(),
      
    
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

######################################



expo_3_60_plot/expo_3_40_plot


```

# Weibull distribution and Reliability 

Definir o conceito de reliability

apresentar a Bathtub theory

introduzir a weibull distribution

exercicios com diferentes shapes

REFAZER DISTRIBUIÇÕES COM A FORMULA
```{r "Weibull distribution", echo=FALSE, message=FALSE, warning=FALSE}
dweibull(x = 10, shape = 1, )

y =dexp(x, rate = 0.05)
x = 0:180


ggplot()+geom_line(aes(x=x, y=y))
```

